Discovery of the first quadruple gravitationally lensed quasar candidate with Pan-STARRS

### 2017-05-26 manuscript.tex
added arxiv manuscript

\author{C. T. Berghea\altaffilmark{1}, G. Nelson\altaffilmark{1}, C. E. Rusu\altaffilmark{2}, C. R. Keeton\altaffilmark{3}, R. P. Dudik\altaffilmark{1}}

\altaffiltext{1}{U.S. Naval Observatory (USNO), 3450 Massachusetts Avenue NW, Washington, DC 20392, USA}
\altaffiltext{2}{Department of Physics, University of California, Davis, 1 Shields Avenue, CA 95616, USA}
\altaffiltext{3}{Rutgers, the State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854, USA}

@@

\keywords{black hole physics --- galaxies: individual (Holmberg IX) --- X-rays: binaries}

@@

Since the serendipitous discovery of the first gravitationally lensed quasar \citep{wal79}, these systems have become powerful probes of astrophysics and cosmology, as illustrated by the wealth of science they have provided over the years. For example, \citep[e.g.,][]{pen06,din17} have studied the coevolution of supermassive black holes and the quasar hosts harboring them up to high redshift. \citet{ogu12,bon17} and others have constrained the cosmological constant and the Hubble constant from samples of lensed quasars. Flux ratio anomalies have revealed luminous satellites or set constraints on dark matter substructure in the lensing galaxies \citep[e.g.,][]{chi05,mck07,fad12}. Ensembles of lenses have revealed the structure of massive galaxies \citep[e.g.,][]{kho00,ogu14}. Other insights have been gained into the quasar accretion disks \citep[e.g.,][]{dai10}, the broad-line regions \citep[e.g.,][]{slu12}, and the black hole spin \citep{reis14}.

These applications all require increasingly large samples of lensed quasars, and in particular quadruply lensed quasars (quads), due to the increased number of modeling constraints they provide. To date, there are only about two dozen known quads over the whole sky. Due to the rare nature of lensed quasars, large scale surveys of sufficient depth and resolution are required to significantly increase the present sample \citep{ogu10}. As such surveys have become available, pioneering searches for lensed quasars, such as the Cosmic Lens All Sky Survey \citep[CLASS;][]{mye03} have been succeeded by the SDSS Quasar Lens Search \citep[SQLS;][]{ogu06}, and more recently by the STRong lensing Insight in the Dark Energy Survey \citep[STRIDES;][]{agn15} and the Survey of Gravitationally-lensed Objects in HSC Imaging \citep[SuGOHI;][]{son17}, amongst others.

@@

@@

We used the stacked images from PS1 in all five filters in order to perform a morphological modeling. We present a close-up color image of the system in Fig.~\ref{lens}, which clearly shows four objects within $\lesssim 3.8\arcsec$ of each other. The three brightest of these (A, B and C) are arranged in an arc-like configuration. The PS1 catalogues identify only the sources A and D, with B and C being blended with A.

We originally used a point spread function (PSF) constructed from stars in the PS1 field of view (FOV), but this left large residuals at the locations of the four objects, significant enough to affect the derived parameters. We suspected that this is due to spatial variation in the PSF across the PS1 FOV, and therefore adopted a technique of fitting the system with an analytical PSF, assumed to not change on the small scale represented by the system. We did this using Hostlens \citep{rus16}, a variant of glafic \citep{ogu10b} which uses $\chi^2$ minimization in order to fit point sources as well as S{\'e}rsic \citep{ser63} profiles convolved with an analytical PSF. The analytical PSF is comprised of two concentric Moffat \citep{mof69} profiles, each one with its own full-width half maximum, ellipticity, orientation, and shape parameter. The two profiles are also characterized by a relative flux of one to the other. We successfully modeled the four objects as point sources convolved with the PSF, therefore showing that they are point-like. Our strategy was to run Hostlens starting from 100 different positions in the parameter space, select the best of the resulting models, and further run 10 Markov Chain Monte Carlo (MCMC) chains using the Metropolis–Hastings algorithm around it. The chains consist of one million steps, with an acceptance rate of $\sim0.3$, and we removed the first 1/5 of these (the ``burn-in'' steps). We also checked that the chains have converged, using the method of \citet{gel95}. We show MCMC-derived uncertainties between the various analytical parameters in Fig.~\ref{mcmc_hostlens}. In order to quantify additional uncertainties inherent in our method due to the non-analytical nature of the ``true'' PSF, we ran 100 simulations where we add noise of similar properties to the real images, on top of the best-fit model. Here, we created the best-fit model by using a PSF constructed from nearby stars, and we ran Hostlens on each of the 100 simulations, using an analytical PSF. As a final estimate of our uncertainties, we used the maximum between the MCMC-derived uncertainties and those from the simulations. We were able to obtain much improved residuals in all bands, and we show these in Fig.~\ref{hostlens}.

We found conspicuous residual in the $i$, $z$ and $y$ bands, just north of the faint point source (Fig.~\ref{hostlens}, middle row). This is consistent with the discovery of a lensing galaxy, in support of the lensing nature of this system. Due to the relatively low signal to noise ratio and the proximity of the point sources, we were unable to fit the morphology of this galaxy with free parameters. The red colors, suggested by the non-detection in the bluer bands, suggest that it is a red, early-type galaxy. As a result, we modeled it with a S{\'e}rsic index of 4, typical of early-type galaxies. We checked that this produces a better fit than a S{\'e}rsic index of 1. For the S{\'e}rsic profile we used an axis ratio of 1 and we fixed the effective radius at $0.5\arcsec$. As the galaxy is most conspicuous in the $i$ and $y$ bands, we consider the most reliable estimate of the relative positions of all objects to be derived from the weighted average of the positions measured in these two bands, where we weight by the inverse of the measured uncertainties. The resulting astrometry, as well as photometry in each band, are shown in Table~1. We have corrected the magnitudes for galactic extinction using the maps of \citet{sch11}.

In addition to the PS1 data, we have also looked for archival data of this system. The system is detected in the Gaia \citep{gaia} catalogues where, similar the the PS1 catalogues, only A and D are deblended by the automatic pipeline. We use the Gaia position for component D as absolute astrometric reference position for this system: 26.792307, 46.511273. The errors for these coordinates are 15.6 and 8.9 mas, respectively [NEEDS REFERENCE].

%PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar to our uncorrected magnitudes: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023. Gaia G magnitudes are 15.17 and 18.13 for A and D, respectively.

The system is also bright in the infrared, with the unresolved magnitudes being: 11.524$\pm$0.022, 10.434$\pm$0.020, 6.769$\pm$0.015, and 4.518$\pm$0.023  in the WISE bands $-$ W1, W2, W3 and W4, respectively. The WISE colors W1$-$W2$=$1.1 and W2$-$W3$=$3.7 match very well those of quasars at relatively low redshift \citep[e.g.][]{wu12}.

Finally, the system is very likely also a radio source. The NRAO VLA Sky Survey (NVSS) images show a source, 2MASX J01471020+4630433, with flux 12.6 mJy at 1.4 GHz. The listed position for this source is only 3 arcsec away from the center of the lens. We are currently in the process of further observing the system in the radio with the Very Large Baseline Array (VLBA). Radio data is useful for the study of lensed quasars, as it provides high-resolution imaging of the source, as well as fluxes not affected by microlensing and extinction.

@@

We aim to fit the photometry measured in Section 2 for the point-like sources and the faint red galaxy using spectral energy distribution (SED) templates. The colors of the point-like images are very similar (Table~1), consistent with being multiple images of the same source. Due to their point-like morphology, we suspect these to be quasars, and we therefore fit them with quasar templates in order to check the quality of the fit and to infer photometric redshifts. Photometric redshifts of quasars are notoriously difficult to constrain, due to the relatively featureless continua \citep{ric09} and the variable equivalent widths of the broad emission lines. We followed two approaches:
\begin{itemize}
\item We used the method of \citet{wu10}, which present quasar colors as a function of redshifts. This is based on Sloan Digital Sky survey (SDSS) colors. PS1 bands are similar but not identical to the SDSS bands and we used the corrections of \citep{fin16} to obtain magnitudes in the SDSS system. Following \citet{wu10}, we minimized ${\chi}^2$ to obtain redshifts for each of the point sources. We plot $\chi^2_{\nu}$ in Figure~\ref{redshift} for each quasar image, and also by fitting all sources together. We notice that the curves are quite consistent with each other, as expected from the similar colors of the sources. In the latter case we obtain the best fit at $z = 0.820^{+0.006}_{-0.007}$. We notice a second minimum at $z \sim2.6$. For both minima the $\chi^2_{\nu}$ is $\lesssim1$, indicative of a good fit. 
\item We fitted quasar templates using LePhare \citep{arn99,ilb06}. We obtained a best-fit redshift of $\sim 2.6-2.8$ for all four point sources, in agreement with the second peak of the method above. The $\chi^2$ is between 13 and 125 for 4 degrees of freedom (5 filters). We find poorer fits (larger $\chi^2$) if we use stellar or galactic SED templates, further supporting that the sources are quasars. In particular, the best-fit galaxy templates correspond to starburst galaxies at low redshift $\lesssim0.1$. At such low redshifts the sources would likely be resolved, not point-like. 
\end{itemize}

We also fitted a galaxy SED to the faint galaxy G, resulting in a best-fit $z = 0.57^{+0.20}_{-0.13}$, obtained for an early-type template. For this, we used the magnitudes measured in the $izy$ bands and assumed detection limits of 21 mag in the $gr$ bands, $\sim$ 2 mag fainter than the rest.

@@

The results of the previous sections show that the source is likely a quadruply imaged quasar, lensed by a foreground early-type galaxy. We subsequently modeled the observed configuration (the relative positions of the objects A,B,C,D and G) using gravlens \citep{kee01}. We chose not to use the observed relative fluxes as constraints, as it is well known that at these wavelengths they usually show large discrepancies due to microlensing, millilensing, intrinsic variability or extinction \citep[e.g.,][]{slu08}. For the mass profile of the lensing galaxy we use a Singular Isothermal Ellipsoid (SIE) with external shear. There are 8 constraints from the 4 image positions and 2 constraints from the galaxy position, for a total of 10. Also, there are 9 free parameters: 2 for the galaxy position, 1 for the Einstein radius, 2 for ellipticity, 2 for shear, and 2 for the source position. The model therefore has 1 degree of freedom.

For the best-fit mass model parameters, we find an Einstein radius $b = 1.932_{-0.011}^{+0.008}$, ellipticity $e_c = 0.163_{-0.070}^{+0.063}$, $e_s = -0.035_{-0.024}^{+0.023}$, and shear $\gamma_c = 0.122_{-0.016}^{+0.015}$, $\gamma_s = 0.065\pm0.005$. Here, we expressed the ellipticity and shear in quasi-cartesian coordinates, rather than polar coordinates, as the distribution of uncertainties is more gaussian. The two coordinate systems are related by $(e_c,\gamma_c) = (e,\gamma)\cos2(\theta_e,\theta_\gamma)$ and $(e_s,\gamma_s) = (e,\gamma)\sin2(\theta_e,\theta_\gamma)$. The uncertainties were determined using MCMC. The $\chi^2$ for this model is 0.8.

@@

In Figure~\ref{model} we show the critical curve and the caustics, as well as the positions of the source and the observed images. The source is crossing a cusp, which results in the production of three bright images in close proximity, and one isolated faint image. This is a classic configuration for quad lenses. The predicted magnitude differences are shown in Table1~1, and a comparison with the observed values are shown in Figure~\ref{modelphot}.

We find that the predicted brightness ranking of the images is consistent with the observations, with image A being brightest, and image D being faintest. However, there are still large discrepancies between the observed and predicted fluxes. For example, the fluxes of images B and C are different, whereas they are expected to be the same for any smooth mass profile of the lensing galaxy. Such ``flux anomalies'' have been widely encountered for cusp lensed quasars, and can be explained by small-scale structure in the lensing galaxy \citep[e.g.,][]{kee03}. We also remark that the colors of the four images are very similar, which implies that differential extinction is negligible. In any case, differential extinction would not explain the flux discrepancy in image D, as this is $\sim1$ mag \emph{brighter} than expected. Since image D is closest to the lensing galaxy, the stellar density is comparatively high at its location, meaning that microlensing due to stars in the lensing galaxy is a plausible explanation for the discrepancy. Secondly, we remark that the time delays between image D and the other images is large, which means that while intrinsic flux variations between A, B and C will be washed out, they could stand out in image D, and therefore be responsible for the discrepancy. Lastly, small-scale substructure (millilensing) could cause contribute to the discrepancy.

%Firstly light can be affected as it propagates through the main lens galaxy, for example by differential extinction.  This possibility is interesting in light of the color differences between images.  Propagation effects tend to be more important at optical wavelengths than at radio wavelengths; this is particularly true of extinction, although there can be scatter broadening at radio wavelengths. Secondly, the light can be affected by stars in the lens galaxy through microlensing.  This mainly affects optical wavelengths. Finally, the light can be affected by dark matter substructure in the lens galaxy through millilensing.  This could affect all wavelengths.

\section{CONCLUSIONS AND FUTURE WORKS}

@@

\begin{itemize}
\item The presence of four point-like images within $\lesssim 3.8\arcsec$ of each other, with similar colors, consistent with quasars templates, but less so with stars and galaxies.
\item The detection of a red galaxy between the point source, such that the relative positions of all five objects are fully consistent with a well-known ``cusp'' configuration. We successfully reproduce this configuration using a SIE + shear model for the lensing galaxy. Further more, the brightness ranking of the point-like images is consistent with the one predicted by the model.
 \item The photometric redshift analysis shows that the galaxy is of elliptical template, typical for lensing galaxies, and of redshift smaller than the point sources. For the inferred redshifts (assuming a source redshift of $\sim2.6$), the velocity dispersion of the lensing galaxy is consistent with those of massive early-type galaxies.
\end{itemize}

@@

We note that this system is very similar to the well studied quad lens RX J1131-1231 \citep{slu03}, in terms of image separation and overall configuration. However, our system appears to have a larger source redshift and brighter images by about 2 magnitudes \citep{slu06}. \citet{slu08} also measure flux discrepancies for RX J1131-1231, and show that these flux discrepancies can be explained by microlensing.

@@

The Pan-STARRS1 Surveys (PS1) and the PS1 public science archive have been made possible through contributions by the Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, the Max-Planck Society and its participating institutes, the Max Planck Institute for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, the Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las Cumbres Observatory Global Telescope Network Incorporated, the National Central University of Taiwan, the Space Telescope Science Institute, the National Aeronautics and Space Administration under Grant No. NNX08AR22G issued through the Planetary Science Division of the NASA Science Mission Directorate, the National Science Foundation Grant No. AST-1238877, the University of Maryland, Eotvos Lorand University (ELTE), the Los Alamos National Laboratory, and the Gordon and Betty Moore Foundation.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.

@@

\bibitem[Chambers et al.(2016)]{cha16} Chambers, K.~C., Magnier, E.~A., Metcalfe, N., et al.\ 2016, arXiv:1612.05560 
\bibitem[Chiba et al.(2005)]{chiba05} Chiba M., Minezaki T., Kashikawa N., Kataza H., Inoue K.~T., 2005, ApJ, 627, 53

@@

\bibitem[Fadely \& Keeton(2012)]{fad12} Fadely R., Keeton C.~R., 2012, MNRAS, 419, 936 
\bibitem[Finkbeiner et al.(2016)]{fin16} Finkbeiner, D.~P., Schlafly, E.~F., Schlegel, D.~J., et al.\ 2016, \apj, 822, 66

@@

\bibitem[Onaka \& al.(2008)]{ona08} Onaka P., Tonry J.~L., Isani S., Lee A., Uyeshiro R., Rae C., Robertson L., Ching G., Proc.\ 2008, \procspie, 7014, 12

\bibitem[Peng et al.(2010)]{peng10} Peng, C.~Y., Ho, L.~C., Impey, C.~D., \& Rix, H.-W.\ 2010, \aj, 139, 2097 
\bibitem[Reis et al.(2014)]{reis14} Reis, R.~C., Reynolds, M.~T., Miller, J.~M., \& Walton, D.~J.\ 2014, \nat, 507, 207 
\bibitem[Richards et al.(2009)]{ric09} Richards G.~T., et al., 2009, ApJS, 180, 67 
\bibitem[Rusu et al.(2016)]{rus16} Rusu C.~E., et al., 2016, MNRAS, 458, 2

>>>>

\author{C. T. Berghea\altaffilmark{1}, George J. Nelson\altaffilmark{1}, C. E. Rusu\altaffilmark{2}, C. R. Keeton\altaffilmark{3}, R. P. Dudik\altaffilmark{1}}

\altaffiltext{1}{U.S. Naval Observatory (USNO), 3450 Massachusetts Avenue NW, Washington, DC 20392, USA}
\altaffiltext{2}{Department of Physics, University of California, Davis, 1 Shields Avenue, CA 95616, USA}
\altaffiltext{3}{Department of Physics \& Astronomy, Rutgers, the State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854, USA}

@@

\keywords{quasars: general  --- gravitational lensing: strong  --- cosmology: observations }

@@

Since the serendipitous discovery of the first gravitationally lensed quasar \citep{wal79}, these systems have become powerful probes of astrophysics and cosmology, as illustrated by the wealth of science they have provided over the years. For example, \citet{peng06,din17} have studied the coevolution of supermassive black holes and the quasar hosts harboring them up to high redshift \citep{cla02,tre10}. \citet{ogu12,bon17} and others have constrained the cosmological constant and the Hubble constant from samples of lensed quasars \citep[see recent review by][]{Tre16}. Flux ratio anomalies have revealed luminous satellites or set constraints on dark matter substructure in the lensing galaxies \citep[e.g.,][]{chiba05,mck07,fad12}. Ensembles of lenses have revealed the structure of massive galaxies \citep[e.g.,][]{koc00,ogu14}. Other insights have been gained into quasar accretion disks \citep[e.g.,][]{dai10}, broad-line regions \citep[e.g.,][]{slu12}, and black hole spin \citep{reis14}.

These applications all require increasingly large samples of lensed quasars, and in particular quadruply lensed quasars (quads), due to the increased number of modeling constraints they provide. To date, there are only about three dozen known quads over the whole sky\footnote{http://masterlens.astro.utah.edu/}. Due to the rare nature of lensed quasars, large scale surveys of sufficient depth and resolution are required to significantly increase the present sample \citep{ogu10}. As such surveys have become available, pioneering searches for lensed quasars, such as the Cosmic Lens All Sky Survey \citep[CLASS;][]{mye03} have been succeeded by the SDSS Quasar Lens Search \citep[SQLS;][]{ogu06}, and more recently by the STRong lensing Insight in the Dark Energy Survey \citep[STRIDES;][]{agn15} and the Survey of Gravitationally-lensed Objects in HSC Imaging \citep[SuGOHI;][]{son17}, amongst others.

@@

The lensed quasar was discovered as part of a variability study of Active Galactic Nuclei selected based on Mid-Infrared characteristics \citep{sec01}, and data from the USNO Robotic Astrometric Telescope \citep{zac15}.

@@

We used the stacked images from PS1 in all five filters in order to perform a morphological modeling. We present a close-up color image of the system in Fig.~\ref{lens}, which clearly shows four objects within $\lesssim 3.8\arcsec$ of each other. The three brightest of these (A, B and C) are arranged in an arc-like configuration. The PS1 catalog identifies only the sources A and D, with B and C being blended with A.

The simple modeling with a point spread function (PSF) constructed from stars in the PS1 field of view (FOV) left large residuals at the locations of the four objects, that were significant enough to affect the derived parameters. We suspect that this is due to spatial variation in the PSF across the PS1 FOV, and have therefore adopted a technique of fitting the system with an analytical PSF, which it is assumed will not change on the small scale represented by the system. We did this using Hostlens \citep{rus16}, a variant of glafic \citep{ogu10b} which uses $\chi^2$ minimization in order to fit point sources as well as S{\'e}rsic \citep{ser63} profiles convolved with an analytical PSF. The analytical PSF is comprised of two concentric Moffat \citep{mof69} profiles, each one with its own full-width half maximum, ellipticity, orientation, and shape parameter. The two profiles are also characterized by a relative flux of one to the other. We successfully modeled the four objects as point sources convolved with the PSF, therefore showing that they are point-like. Our strategy was to run Hostlens starting from 100 different positions in the parameter space, select the best of the resulting models, and further run 10 Markov Chain Monte Carlo (MCMC) chains using the MetropolisÐHastings algorithm around it. The chains consist of one million steps, with an acceptance rate of $\sim0.3$, and we removed the first 1/5 of these (the ``burn-in'' steps). We also checked that the chains have converged, using the method of \citet{gel95}. The MCMC-derived uncertainties between the various analytical parameters are shown in Fig.~\ref{mcmc_hostlens}, and generated using corner.py \citep{for16}. In order to quantify additional uncertainties inherent in our method due to the non-analytical nature of the ``true'' PSF, we ran 100 simulations where we add noise of similar properties to the real images, on top of the best-fit model. Here, we created the best-fit model by using a PSF constructed from nearby stars, and we ran Hostlens on each of the 100 simulations, using an analytical PSF. As a final estimate of our uncertainties, we used the maximum between the MCMC-derived uncertainties and those from the simulations. We were able to obtain much improved residuals in all bands, and we show these in Fig.~\ref{hostlens}.

Fig.~\ref{hostlens} shows conspicuous residuals in the $i$, $z$ and $y$ bands (middle row). These are consistent with the discovery of a lensing galaxy, in support of the lensing nature of this system. Due to the relatively low signal to noise ratio and the proximity of the point sources, we were unable to fit the morphology of this galaxy with free parameters. The red colors, suggested by the non-detection in the bluer bands, suggest that it is a red, early-type galaxy. As a result, we modeled it with a S{\'e}rsic index of 4, typical of early-type galaxies. We checked that this produces a better fit than a S{\'e}rsic index of 1. For the S{\'e}rsic profile we used an axis ratio of 1 and we fixed the effective radius at $0.5\arcsec$. As the galaxy is most conspicuous in the $i$ and $y$ bands, we consider the most reliable estimate of the relative positions of all objects to be derived from the weighted average of the positions measured in these two bands, where we weight by the inverse of the measured uncertainties. The resulting astrometry, as well as photometry in each band, are shown in Table~1. We have corrected the magnitudes for galactic extinction using the maps of \citet{sch11}.

In addition to the PS1 data, we have also looked for archival data of this system. The system is detected in both the Gaia \citep{gaia} DR1 and PS1 catalogs; however, only A and D are deblended by the automatic pipeline. PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar to our uncorrected magnitudes: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023 in the $grizy$ bands, respectively. We use the Gaia position for component D as absolute astrometric reference position for this system: 26.792307, 46.511273. The errors for these coordinates given in the Gaia catalog are 15.6 and 8.9 mas, respectively.

%%PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar to our uncorrected magnitudes: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023. Gaia G magnitudes are 15.17 and 18.13 for A and D, respectively.

The system is also bright in the infrared, with the unresolved magnitudes being: 11.524$\pm$0.022, 10.434$\pm$0.020, 6.769$\pm$0.015, and 4.518$\pm$0.023 in the WISE bands $-$ W1, W2, W3 and W4, respectively. The WISE colors W1$-$W2$=$1.1 and W2$-$W3$=$3.7 match very well those of quasars for redshift below 3.5 \citep[e.g.][]{wu12}.

Finally, the system is very likely also a radio source. The NRAO VLA Sky Survey (NVSS) images show a source, 2MASX J01471020+4630433, with flux 12.6 mJy at 1.4 GHz. The listed position for this source is only $3\arcsec$ away from the center of the lens, which is comparable with NVSS astrometric errors \citep{con98}. We are currently in the process of further observing the system in the radio with the Very Large Baseline Array (VLBA). Radio data is useful for the study of lensed quasars, as it provides high-resolution imaging of the source, as well as fluxes not affected by microlensing and extinction.

@@

We fit the photometry measured in Section 2 for the point-like sources and the faint red galaxy using spectral energy distribution (SED) templates. The colors of the point-like images are very similar (Table~1), consistent with being multiple images of the same source. Due to their point-like morphology, we suspect these to be quasars, and have therefore fit them with quasar templates in order to check the quality of the fit and to infer photometric redshifts. Photometric redshifts of quasars are notoriously difficult to constrain, due to the relatively featureless continua \citep{ric09} and the variable equivalent widths of the broad emission lines. Redshift was measured using two separate methods:
\begin{itemize}
\item The method of \citet{wu10} presents quasar colors as a function of redshifts. This is based on Sloan Digital Sky survey (SDSS) colors. PS1 bands are similar but not identical to the SDSS bands and we used the corrections of \citet{fin16} to obtain magnitudes in the SDSS system. Following \citet{wu10}, we minimized ${\chi}^2$ to obtain redshifts for each of the point sources. We plot $\chi^2_{\nu}$ in Figure~\ref{redshift} for each quasar image, and also by fitting all sources together. We notice that the curves are quite consistent with each other, as expected from the similar colors of the sources. In the latter case we obtain the best fit at $z = 0.820^{+0.006}_{-0.007}$. We notice a second minimum at $z \sim2.6$. For both minima the $\chi^2_{\nu}$ is $\lesssim1$, indicative of a good fit. 
\item LePhare \citep{arn99,ilb06} method uses quasar templates. We obtain the best fit at a redshift of $\sim 2.6-2.8$ for all four point sources.  These results agree with the second peak of the method above. The $\chi^2$ is between 13 and 125 for 4 degrees of freedom (5 filters). We find poorer fits (larger $\chi^2$) when using stellar or galactic SED templates, providing further support that the sources are quasars. In particular, the best-fit galaxy templates correspond to starburst galaxies at low redshift $\lesssim0.1$. At such low redshifts the sources would likely be resolved, not point-like. 
\end{itemize}

The lens galaxy G is very faint but we managed to fit a galaxy SED with an early-type template, resulting in a best-fit $z = 0.57^{+0.20}_{-0.13}$. For this, we used the magnitudes measured in the $izy$ bands and assumed detection limits of 21 mag in the $gr$ bands, $\sim$ 2 mag fainter than the rest.

@@

The results of the previous sections show that the source is likely a quadruply imaged quasar, lensed by a foreground early-type galaxy. We subsequently modeled the observed configuration (the relative positions of the objects A,B,C,D and G) using the lensmodel software from the gravlens package \citep{kee01}. We chose not to use the observed relative fluxes as constraints, as it is well known that at these wavelengths they usually show large discrepancies due to microlensing, millilensing, intrinsic variability or extinction \citep[e.g.,][]{slu08}. For the mass profile of the lensing galaxy we use a Singular Isothermal Ellipsoid (SIE) with external shear. There are 8 constraints from the 4 image positions and 2 constraints from the galaxy position, for a total of 10. Also, there are 9 free parameters: 2 for the galaxy position, 1 for the Einstein radius, 2 for ellipticity, 2 for shear, and 2 for the source position. The model therefore has 1 degree of freedom.

For the best-fit mass model parameters, we find an Einstein radius $b = 1.932_{-0.011}^{+0.008}$, ellipticity $e_c = 0.163_{-0.070}^{+0.063}$, $e_s = -0.035_{-0.024}^{+0.023}$, and shear $\gamma_c = 0.122_{-0.016}^{+0.015}$, $\gamma_s = 0.065\pm0.005$. Here, we expressed the ellipticity and shear in quasi-cartesian coordinates, rather than polar coordinates, as the distribution of uncertainties is more gaussian. The two coordinate systems are related by $(e_c,\gamma_c) = (e,\gamma)\cos2(\theta_e,\theta_\gamma)$ and $(e_s,\gamma_s) = (e,\gamma)\sin2(\theta_e,\theta_\gamma)$. The convergence was again checked using the method of \citet{gel95}. The uncertainties were determined using MCMC as plotted in Fig.~\ref{mcmc_gravlens} with corner.py. The $\chi^2$ for this model is 0.8.

@@

In Figure~\ref{model} we show the critical curve and the caustics, as well as the positions of the source and the observed images. The source is crossing a cusp, which results in the production of three bright images in close proximity, and one isolated faint image. This is a classic configuration for quad lenses. The predicted magnitude differences are shown in Table~1, and a comparison with the observed values are shown in Figure~\ref{modelphot}.

We find that the predicted brightness ranking of the images is consistent with the observations, with image A being brightest, and image D being faintest. However, there are still large discrepancies between the observed and predicted fluxes. For example, the fluxes of images B and C are different, whereas they are expected to be the same for any smooth mass profile of the lensing galaxy. Such ``flux anomalies'' have been widely encountered for cusp lensed quasars, and can be explained by small-scale structure in the lensing galaxy \citep[e.g.,][]{kee03}. We also remark that the colors of the four images are very similar, within $\lesssim 0.2$ mag across all filters. Small chromatic variations between multiple quasar images are known to be caused by intrinsic variability, microlensing or differential extinction \citep{yon08}. In any case, differential extinction would not explain the flux discrepancy in image D, as this is $\sim1$ mag \emph{brighter} than expected. Since image D is closest to the lensing galaxy, the stellar density is comparatively high at its location, meaning that microlensing due to stars in the lensing galaxy is a plausible explanation for the discrepancy. Secondly, we remark that the time delays between image D and the other images is large, which means that while intrinsic flux variations between A, B and C will be washed out, they could stand out in image D, and therefore be responsible for the discrepancy. Lastly, small-scale substructure (millilensing) could cause contribute to the discrepancy.

\section{CONCLUSIONS AND FUTURE WORK}

@@

\begin{itemize}
\item The presence of four point-like images within $\lesssim 3.8\arcsec$ of each other, with similar colors, consistent with quasars templates, but less so with stars and galaxies.
\item The detection of a red galaxy between the point source, such that the relative positions of all five objects are fully consistent with a well-known ``cusp'' configuration. We successfully reproduce this configuration using a SIE + shear model for the lensing galaxy. Furthermore, the brightness ranking of the point-like images is consistent with the one predicted by the model.
 \item The photometric redshift analysis shows that the galaxy is of elliptical template, typical for lensing galaxies, and of redshift smaller than the point sources. For the inferred redshifts (assuming a source redshift of $\sim2.6$), the velocity dispersion of the lensing galaxy is consistent with those of massive early-type galaxies.
\end{itemize}

@@

Finally, we note that this system is very similar to the well-studied quad lens RX J1131-1231 \citep{slu03}, in terms of image separation and overall configuration. However, our system appears to have a larger source redshift and brighter images by about 2 magnitudes \citep{slu06}. \citet{slu08} also measure flux discrepancies for RX J1131-1231, and show that these flux discrepancies can be explained by microlensing.

@@

We would like to thank Arunav Kundu, Shobita Satyapal, Nathan Secrest, Bryan Dorland and Valeri Makarov for useful discussions and help.
The Pan-STARRS1 Surveys (PS1) and the PS1 public science archive have been made possible through contributions by the Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, the Max-Planck Society and its participating institutes, the Max Planck Institute for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, the Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las Cumbres Observatory Global Telescope Network Incorporated, the National Central University of Taiwan, the Space Telescope Science Institute, the National Aeronautics and Space Administration under Grant No. NNX08AR22G issued through the Planetary Science Division of the NASA Science Mission Directorate, the National Science Foundation Grant No. AST-1238877, the University of Maryland, Eotvos Lorand University (ELTE), the Los Alamos National Laboratory, and the Gordon and Betty Moore Foundation.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.

@@

\bibitem[Chambers et al.(2016)]{cha16} Chambers, K.~C., Magnier, E.~A., Metcalfe, N., et al.\ 2016, arXiv:1612.05560 
\bibitem[Chiba et al.(2005)]{chiba05} Chiba, M., Minezaki, T., Kashikawa, N., Kataza, H., \& Inoue, K.~T.\ 2005, \apj, 627, 53

\bibitem[Claeskens \& Surdej(2002)]{cla02} Claeskens, J.-F., \& Surdej, J.\ 2002, \aapr, 10, 263
\bibitem[Condon et al.(1998)]{con98} Condon, J.~J., Cotton, W.~D., Greisen, E.~W., et al.\ 1998, \aj, 115, 1693

@@

\bibitem[Fadely \& Keeton(2012)]{fad12} Fadely R., Keeton C.~R., 2012, MNRAS, 419, 936 
\bibitem[Finch \& Zacharias(2016)]{fin01} Finch, C.~T., \& Zacharias, N.\ 2016, AAS/Division of Dynamical Astronomy Meeting, 47, 203.03 
\bibitem[Finkbeiner et al.(2016)]{fin16} Finkbeiner, D.~P., Schlafly, E.~F., Schlegel, D.~J., et al.\ 2016, \apj, 822, 66
\bibitem[Foreman-Mackey(2016)]{for16} Foreman-Mackey, D. 2016, The Journal of Open Source Software, 24, doi:10.21105/joss.00024

@@

\bibitem[Onaka \& al.(2008)]{ona08} Onaka P., Tonry J.~L., Isani S., Lee A., Uyeshiro R., Rae C., Robertson L., Ching G., Proc.\ 2008, \procspie, 7014, 12
\bibitem[Peng et al.(2006)]{peng06} Peng, C.~Y., Impey, C.~D., Rix, H.-W., et al.\ 2006, \apj, 649, 616 
\bibitem[Peng et al.(2010)]{peng10} Peng, C.~Y., Ho, L.~C., Impey, C.~D., \& Rix, H.-W.\ 2010, \aj, 139, 2097 
\bibitem[Reis et al.(2014)]{reis14} Reis, R.~C., Reynolds, M.~T., Miller, J.~M., \& Walton, D.~J.\ 2014, \nat, 507, 207 
\bibitem[Richards et al.(2009)]{ric09} Richards G.~T., et al., 2009, ApJS, 180, 67 
\bibitem[Rusu et al.(2016)]{rus16} Rusu C.~E., et al., 2016, MNRAS, 458, 2
